---
title: "Car Price Prediction Using Linear Regression"
output: html_document
---

# Car Price Prediction Using Linear Regression

## Overview
### This project applies linear regression modeling to predict car prices using vehicle features. The analysis includes a baseline model, feature selection, advanced feature engineering, and performance evaluation. Model performance is assessed using R² and RMSE metrics.
### This project was completed using R and the `tidymodels` framework as part of a DataCamp project.

## Load Libraries and Data

```{r}
# Load required libraries
library(tidymodels)
library(tidyverse)
library(vip)
library(broom)

# Load training and test datasets
train <- read_csv("train.csv", show_col_types = FALSE)
test <- read_csv("test.csv", show_col_types = FALSE)
```

## Baseline Model: Linear Regression
### We start with a simple linear regression model using all available features, with no transformations. Model performance is evaluated on the test dataset.

```{r}
# Define baseline linear regression model
model <- linear_reg()

# Define recipe with all features
recipe_plain <- recipe(Price ~ ., data = train)

# Define evaluation metrics
metrics <- metric_set(rsq, rmse)

# Fit baseline model and evaluate
original_performance <- 
  workflow() %>%  
  add_model(model) %>%  
  add_recipe(recipe_plain) %>%  
  fit(train) %>%  
  augment(new_data = test) %>%  
  metrics(truth = Price, estimate = .pred) %>%  
  select(.estimate) %>%  
  pull()

# Display performance
print("Baseline model performance:")
print(original_performance)
```

## Feature Engineering and Model Refinement
### We identify low-importance features to remove and apply feature transformations to improve model performance.

```{r}
# Features to remove (low predictive power)
removed_features <- c("Chevy", "Pontiac", "Cruise", "Leather", 
                      "Sound", "Buick", "Saturn", "wagon", "coupe")

# Retain remaining features for documentation
retained_features <- names(train)[!(names(train) %in% c(removed_features, "Price"))]

# Enhanced recipe with transformations
recipe <- 
  recipe(Price ~ ., data = train) %>%  
  step_mutate(Efficiency = Efficiency^2) %>%  
  step_poly(Cylinder) %>%  
  step_interact(~ starts_with("Cylinder"):Efficiency) %>%  
  step_rm(removed_features)

# Workflow with refined recipe
workflow <- 
  workflow() %>%  
  add_model(model) %>%  
  add_recipe(recipe)

# Fit the refined model
fit <- workflow %>% fit(train)
```

## Model Evaluation and Performance Improvement
### We compare the refined model's performance with the baseline using R² and RMSE, and calculate improvement percentages.

```{r}
# Evaluate final model on test data
final_performance <- 
  fit %>%  
  augment(new_data = test) %>%  
  metrics(truth = Price, estimate = .pred) %>%  
  select(.estimate) %>%  
  pull()

# Display final model performance
print("Enhanced model performance:")
print(final_performance)

# Calculate improvements
rsq_improvement <- (final_performance[1] - original_performance[1]) / original_performance[1] * 100
rmse_improvement <- (original_performance[2] - final_performance[2]) / original_performance[2] * 100

# Improvement summary
cat("\nModel improvement:\n")
cat("R² increased by", round(rsq_improvement, 2), "%\n")
cat("RMSE decreased by", round(rmse_improvement, 2), "%\n")

# Print retained features
cat("\nRetained features:", length(retained_features), "total\n")
print(retained_features)
```
